### 2018-06-05 first_simultations.Rmd
Add CWV, two traits cases

---
title: "Simple Test coexistence"
author: "XX"
date: "4 juin 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(cowplot)
```

# Single trait

```{r}
devtools::load_all()
#### Keep stable for all
patches  <- 25   # Number of patches
species  <- 25   # Number of species
time     <- 150	 # Length of model run (generations)
initpop  <- 50   # initial population size
n_traits <- 1    # number of traits
k        <- 1.45 # Niche height
c        <- 10 # Niche breadth

# Competition + Dispersion coefficients
A =  0.0005	# Alpha scalar 
d = 0.0#5 # Dispersal percentage

# Generate environment
env <- 1:patches

composition <- array(NA, dim = c(patches, species, time),
                     dimnames = list(paste0("patches", 1:patches),
                                     paste0("species", 1:species),
                                     paste0("time", 1:time))) #where values = N
names(env) <- dimnames(composition)[[1]]

composition[,,"time1"] <- initpop # N

# Generate traits
traits <- generate_traits(species,
                          rep(min(env), n_traits),
                          rep(max(env), n_traits))

# Single trait competition, single trait growth, single trait both -------------
# Goal: compare between the three above-mentioned cases

## Single Trait Competition
trait_type = c(trait1 = "A")
single_trait_compet = multigen(traits = traits, trait_type = trait_type,
                               env = env, time = time, species = species,
                               patches = patches, composition = composition,
                               A = A, d = d, k = k, c = c)
single_trait_compet_final = ifelse(single_trait_compet$compo[,,time] < 2, 0, single_trait_compet$compo[,,time])

## Single Trait Growth
trait_type = c(trait1 = "R")
single_trait_growth = multigen(traits = traits, trait_type = trait_type,
                               env = env, time = time, species = species,
                               patches = patches, composition = composition,
                               A = A, d = d, k = k, c = c)
single_trait_growth_final = ifelse(single_trait_growth$compo[,,time] < 2, 0,
                                   single_trait_growth$compo[,,time])

## Both
trait_type = c(trait1 = "RA")
single_trait_both = multigen(traits = traits, trait_type = trait_type,
                             env = env, time = time, species = species,
                             patches = patches, composition = composition,
                             A = A, d = d, k = k, c = c)

single_trait_both_final = ifelse(single_trait_both$compo[,,time] < 2, 0,
                                 single_trait_both$compo[,,time])

```

```{r}
p5_comp = plot_patch(single_trait_compet$compo, 5, 150) +
    labs(subtitle = "Competition")
p5_grow = plot_patch(single_trait_growth$compo, 5, 150) +
    labs(subtitle = "Growth")
p5_both = plot_patch(single_trait_both$compo,   5, 150) +
    labs(subtitle = "Both")

p15_comp = plot_patch(single_trait_compet$compo, 15, 150) +
    labs(subtitle = "Competition")
p15_grow = plot_patch(single_trait_growth$compo, 15, 150)  +
    labs(subtitle = "Growth")
p15_both = plot_patch(single_trait_both$compo,   15, 150) +
    labs(subtitle = "Both")

plot_patches = plot_grid(p5_grow, p5_comp, p5_both, p15_grow, p15_comp,
                         p15_both, ncol = 3, nrow = 2)
plot_patches
```

## Compute CWM & CWV

```{r}
compute_moments = function(simulation_routine, trait_df = traits) {
    # CWM computation
    ab_rel = (simulation_routine$compo[,,150] / rowSums(simulation_routine$compo[,,150]))
    cwm_df = ab_rel %*% trait_df
    # Values stored in moments_df
    moments_df = as.data.frame(cwm_df)
    colnames(moments_df) = "cwm"
    moments_df = tibble::rownames_to_column(moments_df, "patch")
    moments_df$env = seq(patches)
    # Add CWV
    cwv_df = (ab_rel %*% trait_df^2) - cwm_df^2
    cwv_df = as.data.frame(cwv_df)
    colnames(cwv_df) = "cwv"
    cwv_df = tibble::rownames_to_column(cwv_df, "patch")
    moments_df = dplyr::left_join(moments_df, cwv_df, by = "patch")

    return(moments_df)
}

both_cw = compute_moments(single_trait_both, traits)
colnames(both_cw) = c("patch", "both_cwm", "env", "both_cwv")
growth_cw = compute_moments(single_trait_growth, traits)
colnames(growth_cw) = c("patch", "growth_cwm", "env", "growth_cwv")
compet_cw = compute_moments(single_trait_compet, traits)
colnames(compet_cw) = c("patch", "compet_cwm", "env", "compet_cwv")

all_cw = merge(both_cw, growth_cw, c("patch", "env"))
all_cw = merge(all_cw, compet_cw, c("patch", "env"))

all_cwm = tidyr::gather(all_cw, "type", "cwm_value", dplyr::contains("cwm"))
all_cwm = all_cwm[, c("patch", "env", "type", "cwm_value")]
all_cwm$type = gsub( "_.*$", "", all_cwm$type)
all_cwv = tidyr::gather(all_cw, "type", "cwv_value", dplyr::contains("cwv"))
all_cwv = all_cwv[, c("patch", "env", "type", "cwv_value")]
all_cwv$type = gsub( "_.*$", "", all_cwv$type)

all_cw = merge(all_cwm, all_cwv, by = c("patch", "env", "type"))

plot_cwm = ggplot(all_cw, aes(env, cwm_value, fill = type)) +
    geom_abline(slope = 1, intercept = 0) +
    geom_point(shape = 21, color = "black", size = 2) +
    labs(x = "Environmental Value",
         y = "Community Weighted Mean",
         subtitle = "Timestep: 150, Trait Both") +
    scale_fill_manual(labels = c("Both", "Competition only", "Growth only"),
                      values = viridis::viridis(3)) +
    theme_bw()

plot_cwv = ggplot(all_cw, aes(env, cwv_value, fill = type)) +
    geom_point(shape = 21, color = "black", size = 2) +
    labs(x = "Environmental Value",
         y = "Community Weighted Variance",
         subtitle = "Timestep: 150, Trait Both") +
    scale_fill_manual(labels = c(both = "Both",
                                 compet = "Competition only",
                                 growth = "Growth only"),
                      values = viridis::viridis(3)) +
    theme_bw()

plot_grid(plot_cwm,
          plot_grid(plot_cwv,
                    ggplot(all_cw[which(all_cw$type %in% c("growth", "both")), ],
                           aes(env, cwv_value, fill = type)) +
                        geom_point(size = 2, shape = 21, color = "black") +
                        labs(x = "Environmental Value",
                             y = "Community Weighted Variance",
                             subtitle = "Timestep: 150, Trait Both") +
                        scale_fill_manual(labels = c(both = "Both",
                                                     growth = "Growth only"),
                                          values = viridis::viridis(3)[c(1, 3)]) +
                        theme_bw(),
                    nrow = 1),
          nrow = 2)

```

```{r}

devtools::load_all()
#### Keep stable for all
patches  <- 25   # Number of patches
species  <- 25   # Number of species
time     <- 150	 # Length of model run (generations)
initpop  <- 50   # initial population size
n_traits <- 1    # number of traits
k        <- 1.45 # Niche height
c        <- 10 # Niche breadth

# Competition + Dispersion coefficients
A =  0.0005	# Alpha scalar 
d = 0.0#5 # Dispersal percentage

# Generate environment
env <- 1:patches

composition <- array(NA, dim = c(patches, species, time),
                     dimnames = list(paste0("patches", 1:patches),
                                     paste0("species", 1:species),
                                     paste0("time", 1:time))) #where values = N
names(env) <- dimnames(composition)[[1]]

composition[,,"time1"] <- initpop # N

# Generate traits
traits <- generate_traits(species,
                          rep(min(env), n_traits),
                          rep(max(env), n_traits))
set.seed(1)
traits <- cbind(traits, trait2 = sample(traits[, 1]))

# Pairs of traits
trait_types = expand.grid(trait1 = c("R", "A", "RA", "none"),
                          trait2 = c("R", "A", "RA", "none"))

# trait_types = as.data.frame(t(combn(c("R", "A", "RA", "none"), 2)))
# colnames(trait_types) <- c("trait1", "trait2")

two_traits <- apply(trait_types, 1,
                    function(x){
                        two_traits_types = multigen(traits = traits, trait_type = x,
                                                    env = env, time = time, species = species,
                                                    patches = patches, composition = composition,
                                                    A = A, d = d, k = k, c = c)
                    })
names(two_traits) <- paste(trait_types[, 1], trait_types[, 2], sep = "_")

plot_patches_list <- lapply(names(two_traits),
                            function(x){
                                plot_patch(two_traits[[x]]$compo, 15, 150) +
                                    labs(subtitle = x)
                            })

plot_grid(plotlist = plot_patches_list,
          ncol = 4, nrow = 4)

# Plot moments
lapply(names(two_traits),
       function(x){
           compute_moments(two_traits[[x]], traits)
       })
both_cw = compute_moments(single_trait_both, traits)
colnames(both_cw) = c("patch", "both_cwm", "env", "both_cwv")

```