### 2018-09-07 solving_A.Rmd
Simple document to explain equations we used to get a value for A

---
title: "Solving Equation to get A value"
author: "Matthias Grenié"
date: "7 septembre 2018"
header-includes:
    - \usepackage{cancel}
    - \usepackage{mathtools}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We start from this equation
$$
N_{t+1, i, x} = \frac{
k \times \exp\left(- \displaystyle\frac{(\text{trait}_i - \text{env}_x)^2}{2\times \text{width}^2}\right) \times N_{t, i, x}}{
1 + A \times \displaystyle\sum_{j = 1,~j \neq i}^{S} N_{t, j, x}  (1 - \delta_{ij}) +
B \times N_{t,i,x}
}
$$
to get realistic value for $A$ and $B$.
We solve $N_{t+1, i, x} = N_{t, i, x}$ with $A = B$ and $1 - \delta_{ij} = 1$ for all $i$ and $j$ and for $\text{trait}_i - \text{env}_x = 0$.
We thus get:

\begin{equation}
    \begin{aligned}
    N_{t+1, i, x} =& N_{t, i, x}\\
    \Leftrightarrow N_{t, i, x} =& \frac{
        k  \times N_{t, i, x}}{
        1 + A \times \displaystyle\sum_{j = 1,~j \neq i}^{S} N_{t, j, x} +
        A \times N_{t,i,x}}\\
    \Leftrightarrow  N_{t, i, x} =& \frac{
        k  \times N_{t, i, x}}{
        1 + A \times \displaystyle\sum_{j = 1}^{S} N_{t, j, x}}\\
    \Leftrightarrow  \bcancel{N_{t, i, x}} =& \frac{
        k  \times \bcancel{N_{t, i, x}}}{
        1 + A \times \displaystyle\sum_{j = 1}^{S} N_{t, j, x}}\\
    \Leftrightarrow  1 =& \frac{k}{
        1 + A \times \displaystyle\sum_{j = 1}^{S} N_{t, j, x}}\\
    \Leftrightarrow  1 + A \times \displaystyle\sum_{j = 1}^{S} N_{t, j, x} &= k\\
    \Leftrightarrow  \Aboxed{A =& \frac{k - 1}{
        \displaystyle\sum_{j = 1}^{S} N_{t, j, x}}}
    \end{aligned}
\end{equation}

We can find a value for A heuristically computing it at the last generation for a number of sites