### 2018-10-09 meta_simulations.Rmd
:sparkles: CWM <-> Env. plots

---
title: "fdcoexist"
author:
    - Pierre Denelle, Matthias Grenié, Cyrille Violle and Caroline M. Tucker
date: "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 10.5, cache = TRUE)

# Packages ---------------------------------------------------------------------
suppressPackageStartupMessages({
    suppressWarnings({
        library("cowplot")
        library("tidyverse")
        library("directlabels")
        library("furrr")
    })
})
# Functions --------------------------------------------------------------------
# Compute a data.frame of number of individuals in function of optimal trait
# for competition and growth
landscape_df = function(multiple_traits_scenario, trait_df, trait_weights,
                        time = 150) {

    growth_traits = subset(trait_weights, growth_weight != 0)$trait
    compet_traits = subset(trait_weights, compet_weight != 0)$trait

    population_df = multiple_traits_scenario$compo[,,time] %>%
        as.data.frame() %>%
        rownames_to_column("patch") %>%  # Add patch names as column
        # Transform to tidy data.frame (define name of columns and column name
        # for values)
        gather("species", "N", -patch)

    growth_df = enframe(rowMeans(trait_df[, growth_traits, drop = FALSE]),
                        "species", "growth_trait")

    compet_df = enframe(abs(rowMeans(trait_df[, compet_traits, drop = FALSE]) -
                                12.5), "species", "compet_dist")

    population_df %>%
        inner_join(growth_df, by = "species") %>%
        inner_join(compet_df, by = "species") %>%
        # Add patch number as a new column and compute distance to optimal value
        mutate(patch_optim = gsub("patches", "", patch) %>%
                   as.numeric(),
               optim_dist = abs(growth_trait - patch_optim))
}

# This function tidy a list of simulation routine to a single data.frame
# suitable to make figures.
tidy_simulation_list = function(simulation_list, set_of_traits, trait_scenarios,
                                time = 150, scenar_names = names(trait_scenarios)) {
    simulation_list %>%
        names() %>%
        # Tidy single simulations
        lapply(tidy_simulation,
               given_simul = simulation_list,
               given_traits = set_of_traits,
               given_trait_scenar = trait_scenarios,
               time = time) %>%
        # Re-add names
        set_names(nm = names(simulation_list)) %>%
        # Join into single data.frame
        bind_rows(.id = "scenario") %>%
        # Compute pre-CWMs
        mutate_at(vars(contains("trait")), funs(cwm = rel_abund * .)) %>%
        mutate(scenario = fct_relevel(scenario, scenar_names))
}

tidy_simulation = function(scenario, given_simul, given_traits,
                           given_trait_scenar, time = 150) {
    landscape_df(given_simul[[scenario]], given_traits,
                 trait_scenarios[[scenario]], time = time) %>%
        inner_join(given_traits %>%
                       as.data.frame() %>%
                       rownames_to_column("species"),
                   by = "species") %>%
        group_by(patch) %>%
        mutate(rel_abund = N/sum(N)) %>%
        ungroup()
}
# Initial parameters -----------------------------------------------------------
suppressMessages(devtools::load_all())

n_patches <- 25   # Number of patches
n_species <- 100  # Number of species
n_gen     <- 150  # Length of model run (generations)
initpop   <- 50   # initial population size
n_traits  <- 3    # number of traits
k         <- 1.15 # Maximum growth rate
env_width <- rep(10, length.out = n_patches)  # Envtal filter strength
var_width <- seq(10, 1, length.out = n_patches) # Varying envtal filter strength
total_runs = 100  # Number of set of traits to test

# Competition + Dispersion coefficients
B = 2e-4 # Scalar for intraspecific competition
A = 2e-4 # Alpha scalar
d = 0.05 # Dispersal percentage

# Generate environment
env <- seq(n_patches)

# Array for simulations
composition <- array(NA, dim = c(n_patches, n_species, n_gen),
                     dimnames = list(paste0("patches", seq(n_patches)),
                                     paste0("species", seq(n_species)),
                                     paste0("time", seq(n_gen))))
names(env) <- dimnames(composition)[[1]]

composition[,, "time1"] <- initpop

# Define scenarios -------------------------------------------------------------
# Scenarios on traits affecting growth and competition
trait_scenarios = list(
    R0A0 = data.frame(trait = paste0("trait", 1:3),
                      growth_weight = c(1, 0, 0),
                      compet_weight = c(0, 0, 1)),
    R0A100 = data.frame(trait = paste0("trait", 1:3),
                        growth_weight = c(1, 0, 0),
                        compet_weight = c(0, 1, 0)),
    R10A90   = data.frame(trait = paste0("trait", 1:3),
                          growth_weight = c(0.90, 0.10, 0),
                          compet_weight = c(0,    0.90, 0.10)),
    R25A75   = data.frame(trait = paste0("trait", 1:3),
                          growth_weight = c(0.75, 0.25, 0),
                          compet_weight = c(0,    0.75, 0.25)),
    R50A50   = data.frame(trait = paste0("trait", 1:3),
                          growth_weight = c(0.5, 0.5, 0),
                          compet_weight = c(0,   0.5, 0.5)),
    R75A25   = data.frame(trait = paste0("trait", 1:3),
                          growth_weight = c(0.25, 0.75, 0),
                          compet_weight = c(0,    0.25, 0.75)),
    R90A10   = data.frame(trait = paste0("trait", 1:3),
                          growth_weight = c(0.10, 0.90, 0),
                          compet_weight = c(0,    0.10, 0.90)),
    R100A0   = data.frame(trait = paste0("trait", 1:3),
                          growth_weight = c(0, 1, 0),
                          compet_weight = c(0, 0, 1)),
    R100A100 = data.frame(trait = paste0("trait", 1:3),
                          growth_weight = c(0, 1, 0),
                          compet_weight = c(0, 1, 0))
)

# Define meta-function to run all analyses with many trait levels --------------

meta_simul = function(seed_number, given_k = k, given_A = A) {
    set.seed(seed_number)
    uncor_traits = generate_cor_traits(n_patches, n_species, n_traits - 1,
                                       cor_coef = 0)
    set.seed(seed_number)
    low_cor_traits = generate_cor_traits(n_patches, n_species, n_traits - 1,
                                         cor_coef = 0.3)

    set.seed(seed_number)
    high_cor_traits = generate_cor_traits(n_patches, n_species, n_traits - 1,
                                          cor_coef = 0.7)
    all_cor = list(uncor    = uncor_traits,
                   low_cor  = low_cor_traits,
                   high_cor = high_cor_traits)

    all_env = list(constant          = env_width,
                   constant_strong   = env_width/2,
                   constant_stronger = env_width/5
                # varying           = var_width
                   )

    all_compet = list(no_compet = list(A = 0, B = B),
                      compet    = list(A = given_A, B = B))

    all_facets = expand.grid(compet_status = names(all_compet),
                             env_width     = names(all_env),
                             cor_level     = names(all_cor),
                             scenario      = names(trait_scenarios))

    apply(all_facets, 1, function(given_row) {
        multigen(traits = all_cor[[given_row[["cor_level"]]]],
                 trait_weights = trait_scenarios[[given_row[["scenario"]]]],
                 env = env,
                 time = n_gen,
                 species = n_species,
                 patches = n_patches,
                 composition = composition,
                 A = all_compet[[given_row[["compet_status"]]]][["A"]],
                 B = all_compet[[given_row[["compet_status"]]]][["B"]],
                 d = d,
                 k = given_k,
                 width = all_env[[given_row[["env_width"]]]]) %>%
            landscape_df(all_cor[[given_row[["cor_level"]]]],
                         trait_scenarios[[given_row[["scenario"]]]],
                         time = 150) %>%
            inner_join(all_cor[[given_row[["cor_level"]]]] %>%
                           as.data.frame() %>%
                           rownames_to_column("species"),
                       by = "species") %>%
            group_by(patch) %>%
            mutate(rel_abund = N/sum(N)) %>%
            ungroup() %>%
            mutate(compet_status = given_row[["compet_status"]],
                   env_width     = given_row[["env_width"]],
                   cor_level     = given_row[["cor_level"]],
                   scenario      = given_row[["scenario"]],
                   seed_number = seed_number) %>%
            select(seed_number, compet_status, env_width, cor_level, scenario,
                   everything())
    }) %>%
        bind_rows() %>%
        filter(N != 0) %>%  # Take out absent species
        mutate_at(vars(contains("trait")), funs(cwm = rel_abund * .)) %>%
        mutate(scenario = fct_relevel(scenario, names(trait_scenarios)),
               patch_optim = as.numeric(gsub("patches", "", patch))) %>%
        group_by(seed_number, compet_status, env_width, cor_level, scenario,
                 patch) %>%
        summarise(trait1_cwm  = sum(trait1_cwm),
                  trait2_cwm  = sum(trait2_cwm),
                  trait3_cwm  = sum(trait3_cwm),
                  trait1_var  = var(trait1),
                  trait2_var  = var(trait2),
                  trait3_var  = var(trait3),
                  patch_optim = unique(patch_optim))
}

meta_simul2 = function(seed_number, given_k = k, given_A = A) {
    set.seed(seed_number)
    uncor_traits = generate_cor_traits(n_patches, n_species, n_traits - 1,
                                       cor_coef = 0)
    set.seed(seed_number)
    low_cor_traits = generate_cor_traits(n_patches, n_species, n_traits - 1,
                                         cor_coef = 0.3)

    set.seed(seed_number)
    high_cor_traits = generate_cor_traits(n_patches, n_species, n_traits - 1,
                                          cor_coef = 0.7)
    all_cor = list(low_cor  = low_cor_traits)

    all_env = list(constant = env_width)

    all_compet = list(compet = list(A = given_A, B = B))

    all_facets = expand.grid(compet_status = names(all_compet),
                             env_width     = names(all_env),
                             cor_level     = names(all_cor),
                             scenario      = names(trait_scenarios))

    apply(all_facets, 1, function(given_row) {
        multigen(traits = all_cor[[given_row[["cor_level"]]]],
                 trait_weights = trait_scenarios[[given_row[["scenario"]]]],
                 env = env,
                 time = n_gen,
                 species = n_species,
                 patches = n_patches,
                 composition = composition,
                 A = all_compet[[given_row[["compet_status"]]]][["A"]],
                 B = all_compet[[given_row[["compet_status"]]]][["B"]],
                 d = d,
                 k = given_k,
                 width = all_env[[given_row[["env_width"]]]]) %>%
            landscape_df(all_cor[[given_row[["cor_level"]]]],
                         trait_scenarios[[given_row[["scenario"]]]],
                         time = 150) %>%
            inner_join(all_cor[[given_row[["cor_level"]]]] %>%
                           as.data.frame() %>%
                           rownames_to_column("species"),
                       by = "species") %>%
            group_by(patch) %>%
            mutate(rel_abund = N/sum(N)) %>%
            ungroup() %>%
            mutate(compet_status = given_row[["compet_status"]],
                   env_width     = given_row[["env_width"]],
                   cor_level     = given_row[["cor_level"]],
                   scenario      = given_row[["scenario"]],
                   seed_number = seed_number) %>%
            select(seed_number, compet_status, env_width, cor_level, scenario,
                   everything())
    }) %>%
        bind_rows() %>%
        filter(N != 0) %>%  # Take out absent species
        mutate_at(vars(contains("trait")), funs(cwm = rel_abund * .)) %>%
        mutate(scenario = fct_relevel(scenario, names(trait_scenarios)),
               patch_optim = as.numeric(gsub("patches", "", patch))) %>%
        group_by(seed_number, compet_status, env_width, cor_level, scenario,
                 patch) %>%
        summarise(trait1_cwm  = sum(trait1_cwm),
                  trait2_cwm  = sum(trait2_cwm),
                  trait3_cwm  = sum(trait3_cwm),
                  trait1_var  = var(trait1),
                  trait2_var  = var(trait2),
                  trait3_var  = var(trait3),
                  patch_optim = unique(patch_optim))
}

meta_simul_patch = function(seed_number, given_k = k, given_A = A) {
    set.seed(seed_number)
    uncor_traits = generate_cor_traits(n_patches, n_species, n_traits - 1,
                                       cor_coef = 0)

    all_cor = list(uncor    = uncor_traits)

    all_env = list(constant          = env_width)

    all_compet = list(compet    = list(A = given_A, B = B))

    all_facets = expand.grid(compet_status = names(all_compet),
                             env_width     = names(all_env),
                             cor_level     = names(all_cor),
                             scenario      = names(trait_scenarios))

    all_simuls = apply(all_facets, 1, function(given_row) {
        multigen(traits = all_cor[[given_row[["cor_level"]]]],
                 trait_weights = trait_scenarios[[given_row[["scenario"]]]],
                 env = env,
                 time = n_gen,
                 species = n_species,
                 patches = n_patches,
                 composition = composition,
                 A = all_compet[[given_row[["compet_status"]]]][["A"]],
                 B = all_compet[[given_row[["compet_status"]]]][["B"]],
                 d = d,
                 k = given_k,
                 width = all_env[[given_row[["env_width"]]]])
    #     %>%
    #         landscape_df(all_cor[[given_row[["cor_level"]]]],
    #                      trait_scenarios[[given_row[["scenario"]]]],
    #                      time = 150) %>%
    #         inner_join(all_cor[[given_row[["cor_level"]]]] %>%
    #                        as.data.frame() %>%
    #                        rownames_to_column("species"),
    #                    by = "species") %>%
    #         group_by(patch) %>%
    #         mutate(rel_abund = N/sum(N)) %>%
    #         ungroup() %>%
    #         mutate(compet_status = given_row[["compet_status"]],
    #                env_width     = given_row[["env_width"]],
    #                cor_level     = given_row[["cor_level"]],
    #                scenario      = given_row[["scenario"]],
    #                seed_number = seed_number) %>%
    #         select(seed_number, compet_status, env_width, cor_level, scenario,
    #                everything())
    })
    # %>%
    #     bind_rows() %>%
    #     filter(N != 0) %>%  # Take out absent species
    #     mutate_at(vars(contains("trait")), funs(cwm = rel_abund * .)) %>%
    #     mutate(scenario = fct_relevel(scenario, names(trait_scenarios)),
    #            patch_optim = as.numeric(gsub("patches", "", patch))) %>%
    #     group_by(seed_number, compet_status, env_width, cor_level, scenario,
    #              patch) %>%
    #     summarise(trait1_cwm  = sum(trait1_cwm),
    #               trait2_cwm  = sum(trait2_cwm),
    #               trait3_cwm  = sum(trait3_cwm),
    #               trait1_var  = var(trait1),
    #               trait2_var  = var(trait2),
    #               trait3_var  = var(trait3),
    #               patch_optim = unique(patch_optim))

    list(simul = all_simuls, trait = uncor_traits)
}
# Generate traits --------------------------------------------------------------
trait_seed = 5
# Make traits with different correlations
set.seed(trait_seed)
uncor_traits = generate_cor_traits(n_patches, n_species, n_traits - 1,
                                   cor_coef = 0)
set.seed(trait_seed)
low_cor_traits = generate_cor_traits(n_patches, n_species, n_traits - 1,
                                     cor_coef = 0.3)

set.seed(trait_seed)
high_cor_traits = generate_cor_traits(n_patches, n_species, n_traits - 1,
                                      cor_coef = 0.7)

# Graphical parameters ---------------------------------------------------------
theme_set(theme_bw(12))

# Graphical functions ----------------------------------------------------------
label_scenar_facets = function(scenar_name) {
    scenar_name %>%
        as.character() %>%
        str_remove("R") %>%
        str_split("A") %>%
        map_chr(~paste0("Growth = ",  .x[1], "%\n",
                        "Compet. = ", .x[2], "%"))
}

facet_scenar = facet_wrap(vars(scenario),
                          labeller = labeller(scenario = label_scenar_facets))

lab_compet_env = labeller(
    compet_status = c(compet    = "Intra + Inter competition",
                      no_compet = "Intra Only"),
    env_width     = c(constant          = "Constant\nWide Env. Filter",
                      constant_strong   = "Constant\nStrong Env. Filter",
                      constant_stronger = "Constant\nVery Strong Env. Filter",
                      varying           = "Increasing\nFilter Strength"))

lab_cor = c(uncor    = "r = 0",
            low_cor  = "r = 0.3",
            high_cor = "r = 0.7")

```

```{r low_k}
# Simulation parameters --------------------------------------------------------
low_k = 1.15
low_k_run_char = paste0("N = ", total_runs, "; A = ", A, "; B = ", B, "; k = ",
                        low_k)

# Actual simulation -----------------------------------------------------------
plan(multiprocess)
tictoc::tic()
low_k_simuls = future_map(seq(total_runs), function(x) {
    devtools::load_all()

    meta_simul(x, given_k = low_k)
}, .progress = TRUE)
tictoc::toc()

# Compute summaries ------------------------------------------------------------
# Show shape of CWM <-> across situations
low_k_simuls %>%
    bind_rows() %>%
    ggplot(aes(patch_optim, trait2_cwm, color = cor_level,
               group = interaction(cor_level, seed_number))) +
    geom_line(aes(group = interaction(cor_level, seed_number)), alpha = 1/25) +
    geom_smooth(aes(group = cor_level)) +
    facet_grid(rows = vars(compet_status, env_width), cols = vars(scenario)) +
    scale_color_viridis_d("Trait\nCorrelations", labels = lab_cor) +
    labs(x = "Environment",
         y = "Trait 2 CWM",
         subtitle = "CWM <-> Env. Relationship across all simulations")

# Compute slope of CWM <-> Environment relationship
low_k_simuls_slope = low_k_simuls %>%
    map(~.x %>%
        gather("cwm_name", "cwm_value", ends_with("cwm")) %>%
        group_by(cwm_name, add = TRUE) %>%
        do(mod_cwm_env = lm(cwm_value ~ patch_optim, data = .)) %>%
        ungroup() %>%
        mutate(mod_glance = lapply(mod_cwm_env, broom::tidy)) %>%
        unnest(mod_glance) %>%
            filter(term == "patch_optim")) %>%
    bind_rows()

# Gather the variance of traits along environment
low_k_simuls_var = low_k_simuls %>%
    map(~.x %>%
            select(-ends_with("cwm"))) %>%
    bind_rows() %>%
    gather("var_name", "var_value", ends_with("var"))

# Figures ----------------------------------------------------------------------
plot_low_k_slope = ggplot(low_k_simuls_slope, aes(scenario, estimate)) +
    geom_boxplot(aes(fill = fct_rev(cor_level)),
                 position = position_dodge(width = 0.9)) +
    facet_grid(rows = vars(env_width, compet_status), cols = vars(cwm_name),
               labeller = lab_compet_env) +
    ylim(0, 1) +
    scale_fill_viridis_d("Trait\nCorrelations", labels = lab_cor) +
    labs(subtitle = paste0("Slope of trait2 CWM ~ environment; ",
                           low_k_run_char),
         x = "Scenario",
         y = "Slope") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "top")
plot_low_k_slope

plot_low_k_var = ggplot(low_k_simuls_var, aes(scenario, var_value)) +
    geom_boxplot(aes(fill = fct_rev(cor_level)),
                 position = position_dodge(width = 0.9)) +
    facet_grid(rows = vars(env_width, compet_status), cols = vars(var_name),
               labeller = lab_compet_env, scales = "free_y") +
    ylim(0, NA) +
    scale_fill_viridis_d("Trait\nCorrelations", labels = lab_cor) +
    labs(subtitle = paste0("Trait Variance across Patches; ", low_k_run_char),
         x = "Scenario",
         y = "Trait Variance among patches") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "top")

plot_low_k_var
```

```{r, eval = FALSE}
high_k = 1.57
high_k_run_char = paste0("N = ", total_runs, "; A = ", A, "; B = ", B, "; k = ",
                        high_k)

plan(multiprocess)
tictoc::tic()

high_k_simuls = future_map(seq(total_runs), function(x) {
    devtools::load_all()

    meta_simul(x, high_k)
}, .progress = TRUE)
tictoc::toc()

high_k_simuls_r2 = high_k_simuls %>%
    map(~.x %>%
        gather("cwm_name", "cwm_value", ends_with("cwm")) %>%
        group_by(cwm_name, add = TRUE) %>%
        do(mod_cwm_env = lm(cwm_value ~ patch_optim, data = .)) %>%
        ungroup() %>%
        mutate(mod_glance = lapply(mod_cwm_env, broom::glance)) %>%
        unnest(mod_glance)) %>%
    bind_rows()

high_k_simuls_var = high_k_simuls %>%
    map(~.x %>%
            select(-ends_with("cwm"))) %>%
    bind_rows() %>%
    gather("var_name", "var_value", ends_with("var"))

plot_high_k_r2 = ggplot(high_k_simuls_r2, aes(scenario, r.squared)) +
    geom_boxplot(aes(fill = fct_rev(cor_level)),
                 position = position_dodge(width = 0.9)) +
    facet_grid(rows = vars(env_width, compet_status), cols = vars(cwm_name),
               labeller = lab_compet_env) +
    ylim(0, NA) +
    scale_fill_viridis_d("Trait\nCorrelations", labels = lab_cor) +
    labs(subtitle = paste0("R2 of trait2 CWM ~ environment; ", high_k_run_char),
         x = "Scenario",
         y = expression("R"^2)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "top")

plot_high_k_r2

plot_high_k_var = ggplot(high_k_simuls_var, aes(scenario, var_value)) +
    geom_boxplot(aes(fill = fct_rev(cor_level)),
                 position = position_dodge(width = 0.9)) +
    facet_grid(rows = vars(env_width, compet_status), cols = vars(var_name),
               labeller = lab_compet_env, scales = "free_y") +
    ylim(0, NA) +
    scale_fill_viridis_d("Trait\nCorrelations", labels = lab_cor) +
    labs(subtitle = paste0("Trait Variance across Patches; ", high_k_run_char),
         x = "Scenario",
         y = "Trait Variance among patches") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "top")

plot_high_k_var
```

```{r varying_A_and_k, eval = FALSE}
list_A = 2*10^-(seq(4, 6))
list_k = 1 + c(0.05, 0.15, 0.25)

plan(multiprocess(workers = availableCores() - 1))
tictoc::tic()

var_k_var_a = cross(list(run_n = seq(total_runs), A = list_A, k = list_k)) %>%
    future_map(function(x) {
    devtools::load_all()

    meta_simul2(x$run_n, x$k, x$A) %>%
        mutate(k = x$k, A = x$A)
}, .progress = TRUE)
tictoc::toc()

var_k_var_a_cwm = var_k_var_a %>%
    bind_rows() %>%
    ungroup() %>%
    nest(-k, -A, -seed_number, -scenario) %>%
    mutate(data = data %>%
               map(~.x %>%
                       gather("cwm_name", "cwm_value", ends_with("cwm")) %>%
                       group_by(cwm_name, add = TRUE) %>%
                       do(mod_cwm_env = lm(cwm_value ~ patch_optim, data = .)) %>%
                       ungroup() %>%
                       mutate(mod_glance = lapply(mod_cwm_env, broom::glance)) %>%
                       unnest(mod_glance))) %>%
    unnest(data) %>%
    filter(cwm_name == "trait2_cwm")

var_k_var_a_var = var_k_var_a %>%
    map(~.x %>%
            select(-ends_with("cwm"))) %>%
    bind_rows() %>%
    gather("var_name", "var_value", ends_with("var")) %>%
    filter(var_name == "trait2_var") %>%
    select(k, A, seed_number, everything())
```

```{r looking_at_slopes, eval = FALSE}
list_A = 2*10^-(seq(4, 6))
list_k = 1 + c(0.05, 0.15, 0.25)

plan(multiprocess(workers = availableCores() - 1))
tictoc::tic()

var_k_var_a = cross(list(run_n = seq(total_runs), A = list_A, k = list_k)) %>%
    future_map(function(x) {
    devtools::load_all()

    meta_simul2(x$run_n, x$k, x$A) %>%
        mutate(k = x$k, A = x$A)
}, .progress = TRUE)
tictoc::toc()

var_k_var_a_cwm = var_k_var_a %>%
    bind_rows() %>%
    ungroup() %>%
    nest(-k, -A, -seed_number, -scenario) %>%
    mutate(data = data %>%
               map(~.x %>%
                       gather("cwm_name", "cwm_value", ends_with("cwm")) %>%
                       group_by(cwm_name, add = TRUE) %>%
                       do(mod_cwm_env = lm(cwm_value ~ patch_optim, data = .)) %>%
                       ungroup() %>%
                       mutate(mod_glance = lapply(mod_cwm_env, broom::tidy)) %>%
                       unnest(mod_glance))) %>%
    unnest(data) %>%
    filter(cwm_name == "trait2_cwm")
# 
# var_k_var_a_var = var_k_var_a %>%
#     map(~.x %>%
#             select(-ends_with("cwm"))) %>%
#     bind_rows() %>%
#     gather("var_name", "var_value", ends_with("var")) %>%
#     filter(var_name == "trait2_var") %>%
#     select(k, A, seed_number, everything())
```

We want to follow the dynamics through time to understand the variation in Rsquare
```{r, eval = FALSE}
patch_k_run_char = paste0("N = ", total_runs, "; A = ", A, "; B = ", B, "; k = ",
                        k)

plan(multiprocess)
tictoc::tic()

patch_k_simuls = future_map(seq(total_runs), function(x) {
    devtools::load_all()

    meta_simul_patch(x, )
}, .progress = TRUE)
tictoc::toc()

patch_k_simuls[[1]][[1]][[1]]$compo[15,,] %>%
    as.data.frame() %>%
    rownames_to_column("species") %>%
    gather("time", "N", -species) %>%
    group_by(time) %>%
    mutate(tot_abund = sum(N)) %>%
    ungroup() %>%
    mutate(rel_abund = N / tot_abund) %>%
    inner_join(patch_k_simuls[[1]]$trait %>%
                   as.data.frame() %>%
                   rownames_to_column("species"),
               by = "species") %>%
    mutate_at(vars(starts_with("trait")), funs(precwm = rel_abund * .)) %>%
    group_by(time) %>%
    summarise_at(vars(ends_with("cwm")), sum) %>%
    mutate(time = as.numeric(gsub("time", "", time))) %>%
    gather("cwm_name", "cwm_value", contains("precwm")) %>%
    ggplot(aes(time, cwm_value)) +
    geom_line() +
    facet_grid(cols = vars(cwm_name))
```